---
title: "miniproject1"
author: "Sammit Bal, Thaddeus Poff, Jason Zhang"
date: "2025-03-05"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Front Matter

##1 Read the dataset into R

```{r}
COD_dataset <- read_excel("./CODGames2_mp.xlsx")
```

## Introduction to the Dataset

**Notes for Our Group:**

1.  Each row represents a game
2.  We're interested in the variables "PrimaryWeapon" (J), "Score" (S),
    "Choice" (J), "Eliminations" (T), "TotalXP" (T)


## Variable 1 (PrimaryWeapon)

For the first variable, we selected the quantitative variable Score of
the numeric type, which ranges in values from 0 to around 8000. There
are no missing observations for the Score variable.

```{r}
# Summary statistics for Score variable
COD_dataset %>%
  dplyr::select(Score) %>%
  summarize(min = min(Score),
  max = max(Score),
  mean = mean(Score),
  median = median(Score),
  sd = sd(Score))
```

**Visualization:**

```{r}
# Histogram of Score variable values to show distribution
ggplot(data = COD_dataset,
  mapping = aes(x = Score)) +
  geom_histogram(binwidth = 500, fill = "blue", color = "black") +
  labs(y = "Number of Matches",
  title = "Histogram of COD Scores")
```

**Analysis:**

Based on this analysis, the scores ranged from 100 to around 7300, with
a majority of values falling between 1500 and 4000. The distribution of
Score is slightly right skewed with a mean of Score 2,945 and a median
of 2,770. The histogram does suggest a few matches in which the player
performed at a high level and scored more than

## Variable 2 (Score)

For the first variable, we selected the quantitative variable Score of
the numeric type, which ranges in values from 0 to around 8000. There
are no missing observations for the Score variable.

```{r}
# Summary statistics for Score variable
COD_dataset %>%
  dplyr::select(Score) %>%
  summarize(min = min(Score),
  max = max(Score),
  mean = mean(Score),
  median = median(Score),
  sd = sd(Score))
```

**Visualization:**

```{r}
# Histogram of Score variable values to show distribution
ggplot(data = COD_dataset,
  mapping = aes(x = Score)) +
  geom_histogram(binwidth = 500, fill = "blue", color = "black") +
  labs(y = "Number of Matches",
  title = "Histogram of COD Scores")
```

**Analysis:**

Based on this analysis, the scores ranged from 100 to around 7300, with
a majority of values falling between 1500 and 4000. The distribution of
Score is slightly right skewed with a mean of Score 2,945 and a median
of 2,770. The histogram does suggest a few matches in which the player
performed at a high level and scored more than

## Variable 3 (Choice)

For the first variable, we selected the quantitative variable Score of
the numeric type, which ranges in values from 0 to around 8000. There
are no missing observations for the Score variable.

```{r}
# Summary statistics for Score variable
COD_dataset %>%
  dplyr::select(Score) %>%
  summarize(min = min(Score),
  max = max(Score),
  mean = mean(Score),
  median = median(Score),
  sd = sd(Score))
```

**Visualization:**

```{r}
# Histogram of Score variable values to show distribution
ggplot(data = COD_dataset,
  mapping = aes(x = Score)) +
  geom_histogram(binwidth = 500, fill = "blue", color = "black") +
  labs(y = "Number of Matches",
  title = "Histogram of COD Scores")
```

**Analysis:**

Based on this analysis, the scores ranged from 100 to around 7300, with
a majority of values falling between 1500 and 4000. The distribution of
Score is slightly right skewed with a mean of Score 2,945 and a median
of 2,770. The histogram does suggest a few matches in which the player
performed at a high level and scored more than

## Variable 4 (Eliminations) T

For the first variable, we selected the quantitative variable Score of
the numeric type, which ranges in values from 0 to around 8000. There
are no missing observations for the Score variable.

```{r}
# Summary statistics for Score variable
COD_dataset %>%
  dplyr::select(Score) %>%
  summarize(min = min(Score),
  max = max(Score),
  mean = mean(Score),
  median = median(Score),
  sd = sd(Score))
```

**Visualization:**

```{r}
# Histogram of Score variable values to show distribution
ggplot(data = COD_dataset,
  mapping = aes(x = Score)) +
  geom_histogram(binwidth = 500, fill = "blue", color = "black") +
  labs(y = "Number of Matches",
  title = "Histogram of COD Scores")
```

**Analysis:**

Based on this analysis, the scores ranged from 100 to around 7300, with
a majority of values falling between 1500 and 4000. The distribution of
Score is slightly right skewed with a mean of Score 2,945 and a median
of 2,770. The histogram does suggest a few matches in which the player
performed at a high level and scored more than

## Variable 5 (TotalXP) T

For the first variable, we selected the quantitative variable Score of
the numeric type, which ranges in values from 0 to around 8000. There
are no missing observations for the Score variable.

```{r}
# Summary statistics for Score variable
COD_dataset %>%
  dplyr::select(Score) %>%
  summarize(min = min(Score),
  max = max(Score),
  mean = mean(Score),
  median = median(Score),
  sd = sd(Score))
```

**Visualization:**

```{r}
# Histogram of Score variable values to show distribution
ggplot(data = COD_dataset,
  mapping = aes(x = Score)) +
  geom_histogram(binwidth = 500, fill = "blue", color = "black") +
  labs(y = "Number of Matches",
  title = "Histogram of COD Scores")
```

**Analysis:**

Based on this analysis, the scores ranged from 100 to around 7300, with
a majority of values falling between 1500 and 4000. The distribution of
Score is slightly right skewed with a mean of Score 2,945 and a median
of 2,770. The histogram does suggest a few matches in which the player
performed at a high level and scored more than

## 3. Begin a list of questions that you have about the dataset. Suppose you are working on a project and your manager gives you this dataset. Write 3 questions that you would ask your manager.

1.  

2.  

3.  

• You have not been given any information about what the variables
represent. So, if you are unsure about the meaning of any of the
variables, write your questions. • You have not been given any
information about how the data were collected. If you have questions
about the data collection, write them down. • If you have any other
questions, write them down. • NOTE: Even if you are familiar with the
game and think that you understand the meaning of the variables, you are
still expected to write 3 potential questions that you could ask.

## 4. Asking a question that can be answered using the dataset and data visualization.

Part 1 – You will answer my question: Is the player’s performance, as
quantified by the amount of experience points gained (TotalXP variable)
changing over time? To answer this, you should find a way to extract the
month from the Date variable (we have not covered working with dates in
STAT 380) o The lubridate package (part of tidyverse) is useful for
working with dates o Once you have the months, create side-by-side
boxplots showing the TotalXP based on the Month, where the months are
named (i.e., month should be June, not the number 6). The months should
be in chronological order, not alphabetical order o Based on the plot,
answer the question. If you have knowledge about these types of games,
you can also mention other variables that could affect TotalXP but were
not included in the question.

# Part 2 – You will write your own question, create a data visualization that can help answer your question, and write an answer to your question.

The question/visualization should explore the relationship between at
least 2 of the variables in the dataset. Be sure you include both the
question and the answer to your question based on the visualization you
have created.
