---
title: "miniproject1"
author: "Sammit Bal, Thaddeus Poff, Jason Zhang"
date: "2025-03-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Front Matter
  
```{r}
COD_dataset <- read_excel("./CODGames2_mp.xlsx")
```

## Introduction to the Dataset
**Notes for Our Group:**

1. Each row represents a game
2. We're interested in the variables "PrimaryWeapon" (J), "Score" (S), "Choice" (J), "Eliminations" (T), "TotalXP" (T)

**Description:**

- In the game, 1 point is awarded per second. 
- Points for the round, game, and match will only be awarded to teams and not to the individual players of the team. 
- Match points are awarded to the team who won the best of five game series. 
- Game points are awarded to the winning team at the conclusion of each game.

## Variable 1 (PrimaryWeapon)
For the first variable, we selected the quantitative variable Score of the numeric type, which ranges in values from
0 to around 8000. There are no missing observations for the Score variable.

```{r}
# Summary statistics for Score variable
COD_dataset %>%
  dplyr::select(Score) %>%
  summarize(min = min(Score),
  max = max(Score),
  mean = mean(Score),
  median = median(Score),
  sd = sd(Score))
```

**Visualization:**

```{r}
# Histogram of Score variable values to show distribution
ggplot(data = COD_dataset,
  mapping = aes(x = Score)) +
  geom_histogram(binwidth = 500, fill = "blue", color = "black") +
  labs(y = "Number of Matches",
  title = "Histogram of COD Scores")
```

**Analysis:**

Based on this analysis, the scores ranged from 100 to around 7300, with a majority of values falling between 1500 and 4000. The distribution of Score is slightly right skewed with a mean of Score 2,945 and a median of 2,770. The histogram does suggest a few matches in which the player performed at a high level and scored more than


## Variable 2 (Score)
For the first variable, we selected the quantitative variable Score of the numeric type, which ranges in values from
0 to around 8000. There are no missing observations for the Score variable.

```{r}
# Summary statistics for Score variable
COD_dataset %>%
  dplyr::select(Score) %>%
  summarize(min = min(Score),
  max = max(Score),
  mean = mean(Score),
  median = median(Score),
  sd = sd(Score))
```

**Visualization:**

```{r}
# Histogram of Score variable values to show distribution
ggplot(data = COD_dataset,
  mapping = aes(x = Score)) +
  geom_histogram(binwidth = 500, fill = "blue", color = "black") +
  labs(y = "Number of Matches",
  title = "Histogram of COD Scores")
```

**Analysis:**

Based on this analysis, the scores ranged from 100 to around 7300, with a majority of values falling between 1500 and 4000. The distribution of Score is slightly right skewed with a mean of Score 2,945 and a median of 2,770. The histogram does suggest a few matches in which the player performed at a high level and scored more than

## Variable 3 (Choice)
For the first variable, we selected the quantitative variable Score of the numeric type, which ranges in values from
0 to around 8000. There are no missing observations for the Score variable.

```{r}
# Summary statistics for Score variable
COD_dataset %>%
  dplyr::select(Score) %>%
  summarize(min = min(Score),
  max = max(Score),
  mean = mean(Score),
  median = median(Score),
  sd = sd(Score))
```

**Visualization:**

```{r}
# Histogram of Score variable values to show distribution
ggplot(data = COD_dataset,
  mapping = aes(x = Score)) +
  geom_histogram(binwidth = 500, fill = "blue", color = "black") +
  labs(y = "Number of Matches",
  title = "Histogram of COD Scores")
```

**Analysis:**

Based on this analysis, the scores ranged from 100 to around 7300, with a majority of values falling between 1500 and 4000. The distribution of Score is slightly right skewed with a mean of Score 2,945 and a median of 2,770. The histogram does suggest a few matches in which the player performed at a high level and scored more than

## Variable 4 (Eliminations) T
For the first variable, we selected the quantitative variable Score of the numeric type, which ranges in values from
0 to around 8000. There are no missing observations for the Score variable.

```{r}
# Summary statistics for Score variable
COD_dataset %>%
  dplyr::select(Score) %>%
  summarize(min = min(Score),
  max = max(Score),
  mean = mean(Score),
  median = median(Score),
  sd = sd(Score))
```

**Visualization:**

```{r}
# Histogram of Score variable values to show distribution
ggplot(data = COD_dataset,
  mapping = aes(x = Score)) +
  geom_histogram(binwidth = 500, fill = "blue", color = "black") +
  labs(y = "Number of Matches",
  title = "Histogram of COD Scores")
```

**Analysis:**

Based on this analysis, the scores ranged from 100 to around 7300, with a majority of values falling between 1500 and 4000. The distribution of Score is slightly right skewed with a mean of Score 2,945 and a median of 2,770. The histogram does suggest a few matches in which the player performed at a high level and scored more than


## Variable 5 (TotalXP) T
For the first variable, we selected the quantitative variable Score of the numeric type, which ranges in values from
0 to around 8000. There are no missing observations for the Score variable.

```{r}
# Summary statistics for Score variable
COD_dataset %>%
  dplyr::select(Score) %>%
  summarize(min = min(Score),
  max = max(Score),
  mean = mean(Score),
  median = median(Score),
  sd = sd(Score))
```

**Visualization:**

```{r}
# Histogram of Score variable values to show distribution
ggplot(data = COD_dataset,
  mapping = aes(x = Score)) +
  geom_histogram(binwidth = 500, fill = "blue", color = "black") +
  labs(y = "Number of Matches",
  title = "Histogram of COD Scores")
```

**Analysis:**

Based on this analysis, the scores ranged from 100 to around 7300, with a majority of values falling between 1500 and 4000. The distribution of Score is slightly right skewed with a mean of Score 2,945 and a median of 2,770. The histogram does suggest a few matches in which the player performed at a high level and scored more than


